#Grazer assaay
```{r}
library(csvread)
setwd("C:/Users/bmbug/OneDrive/Documents/MLML recovered/Bull Kelp Project/data/2025")
getwd()
GA = read.csv("Grazer assay surveys.csv", header =TRUE)

library(dplyr)
library(tidyr)
GA = GA %>%
  mutate_at(c(11:14), as.numeric)%>%
  mutate(DATE = as.Date(DATE, format = "%m/%d/%Y"))%>%
  mutate(TIME = as.character(TIME))%>%
  mutate(SEA.URCHIN = PURPLE.SEA.URCHIN + RED.SEA.URCHIN)%>%
  mutate(Treatment = paste(DRIFT.ADDED, ARKEV, sep = " & "))
str(GA)
```

  
#urchin means
```{r}
library(tidyr)
library(dplyr)

overall_means <- GA %>%
  summarise(
    mean_purple = mean(PURPLE.SEA.URCHIN, na.rm = TRUE),
    mean_red    = mean(RED.SEA.URCHIN, na.rm = TRUE),
    mean_urchin = mean(SEA.URCHIN, na.rm = TRUE)
  )

urchin_summary = GA %>%
  group_by(DATE, ARKEV, Treatment)

#calculate densities by treatment
urchin_den_arkev = urchin_summary %>%
  group_by(Treatment)%>%
  summarise(mean_pu_den = mean(PURPLE.SEA.URCHIN, na.rm = TRUE),
            se_pu_den = sd(PURPLE.SEA.URCHIN, na.rm = TRUE)/sqrt(n()),
            mean_ru_den = mean(RED.SEA.URCHIN, na.rm = TRUE),
            se_ru_den = sd(RED.SEA.URCHIN, na.rm = TRUE)/sqrt(n()),
            mean_urchin_den = mean(SEA.URCHIN, na.rm = TRUE),
            se_urchin_den = sd(SEA.URCHIN, na.rm = TRUE)/ sqrt(n()),
            n= n()) %>%
  ungroup()
```


#Density by arkev
```{r}
library(ggplot2)

ggplot(urchin_den_arkev, aes(x = Treatment, y = mean_urchin_den)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black") +
  geom_errorbar(
    aes(ymin = mean_urchin_den - se_urchin_den, ymax = mean_urchin_den + se_urchin_den),
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  labs(
    title = "Sea Urchin Density by Treatment",
    x = "Treatment",
    y = "Mean Sea Urchin Density (#/m²)",
    fill = "Drift Added"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold", angle = 25, hjust = 1, vjust=1),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 15, face = "bold"), 
    title = element_text(size = 15, face = "bold"),
    legend.title = element_text(size=12, face= "bold"),
    legend.text = element_text(size=12, face= "bold"))
```

```{r}
urchin_data <- GA %>%
  filter(!is.na(SEA.URCHIN)) %>%
  mutate(Treatment = as.factor(Treatment))

urchin_data %>%
  group_by(Treatment) %>%
  summarise(p_value = shapiro.test(SEA.URCHIN)$p.value)

library(car)
leveneTest(SEA.URCHIN ~ Treatment, data = urchin_data)
```

#urchin density by treatment
```{r}
kruskal.test(SEA.URCHIN ~ Treatment, data = urchin_data)

# Post-hoc pairwise Wilcoxon tests
pairwise.wilcox.test(urchin_data$SEA.URCHIN, urchin_data$Treatment,
                     p.adjust.method = "BH")
```

```{r}
urchin_aov <- aov(SEA.URCHIN ~ Treatment, data = urchin_data)
summary(urchin_aov)

# Post-hoc pairwise differences
urchin_tuk <- TukeyHSD(urchin_aov)
urchin_tuk
```

```{r}
library(multcompView)
library(ggplot2)

# Tukey HSD and letters
urchin_tuk <- TukeyHSD(urchin_aov)

# Get letters (compact letter display)
urchin_cld <- multcompLetters4(urchin_aov, urchin_tuk)

# Extract only the letters as a named vector
letters_vec <- urchin_cld$Treatment$Letters

# Convert to a tidy data frame
letters_df <- data.frame(
  Treatment = names(letters_vec),
  TukeyLetters = letters_vec,
  stringsAsFactors = FALSE
)

# Merge with your summary
urchin_den_arkev <- urchin_den_arkev %>%
  left_join(letters_df, by = "Treatment")

# Plot
ggplot(urchin_den_arkev, aes(x = Treatment, y = mean_urchin_den)) +
  geom_bar(stat = "identity", fill = "gray30", color = "black") +
  geom_errorbar(aes(ymin = mean_urchin_den - se_urchin_den,
                    ymax = mean_urchin_den + se_urchin_den),
                width = 0.2) +
  geom_text(aes(label = TukeyLetters), vjust = -1.25, size = 6, fontface = "bold") +
  labs(
    title = "Sea Urchin Density by Treatment",
    x = "Treatment",
    y = "Mean Urchin Density (#/m²)"
  ) +
  theme_classic(base_size = 14)
```


#percent change in urchin density
```{r}
urchin_means <- GA %>%
  filter(TIME %in% c(0, 24)) %>%
  group_by(Month, Treatment, TIME) %>%
  summarise(
    mean_purple = mean(PURPLE.SEA.URCHIN, na.rm = TRUE),
    se_purple = sd(PURPLE.SEA.URCHIN, na.rm = TRUE)/sqrt(n()),
    mean_red = mean(RED.SEA.URCHIN, na.rm = TRUE),
    se_red  = sd(RED.SEA.URCHIN, na.rm = TRUE)/ sqrt(n()),
    mean_urchin = mean(SEA.URCHIN, na.rm = TRUE),
    se_urchin = sd(SEA.URCHIN, na.rm = TRUE)/ sqrt(n()),
    .groups = "drop"
  )%>%
  pivot_wider(
    names_from = TIME,
    values_from = c(mean_purple, mean_red, se_purple, se_red, mean_urchin, se_urchin),
    names_prefix = "t"
  ) %>%
  mutate(
    purple_change = ifelse(mean_purple_t0 == 0, NA, ((mean_purple_t24 - mean_purple_t0)/mean_purple_t0)*100),
    red_change    = ifelse(mean_red_t0 == 0, NA, ((mean_red_t24 - mean_red_t0)/mean_red_t0)*100),
    urchin_change = ifelse(mean_urchin_t0 == 0, NA, ((mean_urchin_t24 - mean_urchin_t0)/mean_urchin_t0)*100)
  )

```

```{r}
#long format
urchin_long <- urchin_means %>%
  pivot_longer(
    cols = c(purple_change, red_change, urchin_change),
    names_to = "Urchin_type",
    values_to = "Percent_change"
  )

#bar plot
ggplot(urchin_long, aes(x = Treatment, y = Percent_change, fill = Urchin_type)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("purple_change" = "purple", "red_change" = "red","urchin_change" = "darkgray")) +
  facet_wrap(~Month) +
  labs(
    y = "Percent Change (0 → 24 hr)",
    x = "Treatment",
    fill = "Urchin Type"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", angle= 45, hjust = 1, vjust= 1),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.position = "right"
  )
```

#log transformed change in urchin densities
```{r}
eps <- 0.01 

urchin_means_eps <- urchin_means %>%
  mutate(
    # log-transformed fold change: log10((t24 + eps) / (t0 + eps))
    log_purple_change = log10((mean_purple_t24 + eps) / (mean_purple_t0 + eps)),
    log_red_change = log10((mean_red_t24 + eps) / (mean_red_t0 + eps)),
    log_urchin_change = log10((mean_urchin_t24 +eps) / (mean_urchin_t0 + eps))
  )

log_urchin_means_summary <- urchin_means_eps %>%
  group_by(Month, Treatment) %>%
  summarise(
    mean_log_purple = mean(log_purple_change, na.rm = TRUE),
    se_log_purple = sd(log_purple_change, na.rm = TRUE) / sqrt(sum(!is.na(log_purple_change))),
    mean_log_red = mean(log_red_change, na.rm = TRUE),
    se_log_red = sd(log_red_change, na.rm = TRUE) / sqrt(sum(!is.na(log_red_change))),
    mean_log_urchin = mean(log_urchin_change, na.rm = TRUE),
    se_log_urhcin = sd(log_urchin_change, na.rm= TRUE) / sqrt(sum(!is.na(log_urchin_change))),
    .groups = "drop"
  )

library(tidyr)

log_urchin_long <- log_urchin_means_summary %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "urchin_type",
    values_to = "mean_change"
  ) %>%
  mutate(
    se_change = case_when(
      urchin_type == "mean_log_purple" ~ se_log_purple,
      urchin_type == "mean_log_red" ~ se_log_red,
      urchin_type == "mean_log_urchin" ~ se_log_urhcin
    ),
    urchin_type = case_when(
      urchin_type == "mean_log_purple" ~ "Purple Sea Urchin",
      urchin_type == "mean_log_red" ~ "Red Sea Urchin",
      urchin_type == "mean_log_urchin" ~ "Sea Urchin"
    )
  )
```

```{r}
library(ggplot2)

ggplot(log_urchin_long, aes(x = Treatment, y = mean_change, fill = urchin_type)) +
  geom_col(position = position_dodge(width = 0.8), color = "black", width = 0.7) +
  geom_errorbar(
    aes(ymin = mean_change - se_change, ymax = mean_change + se_change),
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  facet_wrap(~Month, nrow = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +  # 0 = no change
  labs(
    title = "Change in Urchin Density (0 → 24 hr)",
    x = "Treatment",
    y = "Log10 Change in Densities",
    fill = "Urchin Type"
  ) + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("Purple Sea Urchin" = "purple", "Red Sea Urchin" = "red")) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", angle= 45, hjust = 1, vjust= 1),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.position = "right"
  )
  
```


#change in density by quad
```{r}
#analyze by quads
eps <- 0.01  # to avoid log(0)

quad_change <- GA %>%
  filter(TIME %in% c(0, 24)) %>%
  group_by(Month, ARKEV, DRIFT.ADDED, Treatment, QUAD, TIME) %>%
  summarise(
    mean_urchin = mean(SEA.URCHIN, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = TIME,
    values_from = mean_urchin,
    names_prefix = "t"
  ) %>%
  mutate(
    # log10 fold change from 0 to 24 hr
    log_urchin_change = log10((t24 + eps) / (t0 + eps)),
    # extract numeric distance from QUAD (BAR = 0)
    Distance_m = case_when(
      QUAD == "BAR" ~ 0,
      TRUE ~ as.numeric(str_extract(QUAD, "\\d+"))
    )
  )

str(quad_change)

```



library(ggplot2)

ggplot(quad_change, aes(x = Distance_m, y = log_urchin_change, color = Treatment)) +
  geom_point(position = position_jitter(width = 0.2, height = 0), size = 3) +
  geom_line(aes(group = QUAD), alpha = 0.5) +
  facet_wrap(~Month, nrow = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Quadrat-level Log Fold Change in Purple Urchin Density",
    x = "Distance from BAR (m)",
    y = "Log10 Fold Change (t24 / t0)",
    color = "Treatment"
  ) +
  theme_classic(base_size = 14)

ggplot(quad_change, aes(x = factor(Distance_m), y = log_urchin_change, fill = Treatment)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  facet_wrap(~Month, nrow = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Change in Purple Urchin Density",
    x = "Distance from GA BAR (m)",
    y = "Log10 Fold Change",
    fill = "Treatment"
  ) +
  theme_classic(base_size = 14)




eps <- 0.01

quad_summary <- quad_change %>%
  # ensure BAR = 0
  mutate(
     Distance_m = case_when(
      QUAD == "BAR" ~ 0,
      TRUE ~ as.numeric(str_extract(QUAD, "\\d+"))
    ),
    # log fold change
    log_purple_change = log10((mean_purple_t24 + eps) / (mean_purple_t0 + eps)),
    log_red_change = log10((mean_red_t24 + eps) / (mean_red_t0 + eps)),
    log_urchin_change = log10((mean_urchin_t24 + eps) / (mean_urchin_t0 + eps)),
  ) %>%
  group_by(Month, DRIFT.ADDED, Distance_m) %>%
  summarise(
    mean_purple = mean(log_purple_change, na.rm = TRUE),
    se_purple   = sd(log_purple_change, na.rm = TRUE)/sqrt(sum(!is.na(log_purple_change))),
    mean_red    = mean(log_red_change, na.rm = TRUE),
    se_red      = sd(log_red_change, na.rm = TRUE)/sqrt(sum(!is.na(log_red_change))),
    mean_urchin = mean(log_urchin_change, na.rm = TRUE),
    se_urchin = sd(log_urchin_change, na.rm= TRUE) / sqrt(sum(!is.na(log_urchin_change))),
    .groups = "drop"
  ) %>%
  # pivot longer for combined plotting
  pivot_longer(
    cols = c(mean_purple, mean_red, mean_urchin),
    names_to = "Urchin_Type",
    values_to = "Mean_LogChange"
  ) %>%
  mutate(
    SE = case_when(
      Urchin_Type == "mean_purple" ~ se_purple,
      Urchin_Type == "mean_red"    ~ se_red,
      Urchin_Type == "mean_urchin" ~ se_urchin
    ),
    Urchin_Type = case_when(
      Urchin_Type == "mean_purple" ~ "Purple Sea Urchin",
      Urchin_Type == "mean_red"    ~ "Red Sea Urchin",
      Urchin_Type == "mean_urchin" ~ "Sea Urchin"
    )
  )


library(ggplot2)

ggplot(quad_summary, aes(x = Distance_m, y = Mean_LogChange, color = Urchin_Type, shape = DRIFT.ADDED)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = Mean_LogChange - SE, ymax = Mean_LogChange + SE),
                width = 0.2, position = position_dodge(width = 0.3)) +
  facet_wrap(~Month, nrow = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Change in Urchin Density ",
    x = "Distance from GA BAR (m)",
    y = "24 HR Change (log10)",
    color = "Urchin Type"
  ) +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Purple Sea Urchin" = "purple", "Red Sea Urchin" = "red", "Sea Urchin" = "black"))

ggplot(distance_summary, aes(x = Distance_m, y = mean, color = Urchin_type, shape = Treatment)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
              position = position_dodge(width = 0.5), width = 0.5) +
  geom_line(position = position_dodge(width = 0.5), width = 0.5) +
  scale_color_manual(values = c("purple", "red", "black")) +
  labs(
    x = "Distance from GA Bar (m)",
    y = "Log10 Fold Change (24 hr / 0 hr)",
    color = "Urchin Type"
  ) +
   theme_classic(base_size = 14)


```{r}
#combine urchin densities, connect points, change legend to focus on drift

library(dplyr)

distance_summary <- quad_change %>%
  group_by(Treatment, DRIFT.ADDED, Distance_m) %>%
  summarise(
    mean_log_urchin = mean(log_urchin_change, na.rm = TRUE),
    se_log_urchin   = sd(log_urchin_change, na.rm = TRUE) / sqrt(sum(!is.na(log_urchin_change))),
    .groups = "drop"
  )

ggplot(distance_summary, aes(x = Distance_m, y = mean_log_urchin, color = Treatment)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_log_urchin - se_log_urchin,
                    ymax = mean_log_urchin + se_log_urchin),
                width = 0.4, size = 1) +
  geom_line(aes(group = Treatment), size = 1) +
  labs(
    title = "Change in Total Sea Urchin Density (# per m²)",
    x = "Distance from GA Bar (m)",
    y = "Log10 Fold Change (24 hr / 0 hr)",
    color = "Treatment"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 15, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
```

```{r}
# Make sure Treatment is a factor
quad_change <- quad_change %>%
  mutate(Treatment = factor(Treatment),
         Distance_m = factor(Distance_m))


within_treatment_tests <- quad_change %>%
  group_by(Treatment) %>%
  do({
    model <- aov(log_urchin_change ~ Distance_m, data = .)
    data.frame(
      Treatment = unique(.$Treatment),
      p_value = summary(model)[[1]][["Pr(>F)"]][1]
    )
  })

within_treatment_tests
```


```{r}
library(broom)

tukey_within_treatment <- quad_change %>%
  group_by(Treatment) %>%
  do({
    model <- aov(log_urchin_change ~ Distance_m, data = .)
    if(length(unique(.$Distance_m[!is.na(.$log_urchin_change)])) > 1) {
      broom::tidy(TukeyHSD(model))
    } else {
      tibble(term = "Distance_m", contrast = NA, p.adj = NA)
    }
  })

tukey_within_treatment

```

```{r}
within_distance_tests <- quad_change %>%
  group_by(Distance_m) %>%
  do({
    model <- aov(log_urchin_change ~ Treatment, data = .)
    data.frame(
      Distance_m = unique(.$Distance_m),
      p_value = summary(model)[[1]][["Pr(>F)"]][1]
    )
  })

within_distance_tests
```

```{r}
tukey_within_distance <- quad_change %>%
  group_by(Distance_m) %>%
  do({
    model <- aov(log_urchin_change ~ Treatment, data = .)
    broom::tidy(TukeyHSD(model))
  })

tukey_within_distance

```

```{r}
urchin_change_aov <- aov(log_urchin_change ~ Treatment + Distance_m, data = quad_change)

summary(urchin_change_aov)
TukeyHSD(urchin_change_aov)
```

```{r}
library(multcompView)
library(dplyr)

letters_within_treatment <- quad_change %>%
  group_by(Treatment) %>%
  do({
    model <- aov(log_urchin_change ~ Distance_m, data = .)
    tukey <- TukeyHSD(model)
    cld <- multcompLetters4(model, tukey)
    data.frame(
      Distance_m = names(cld$Distance_m$Letters),
      Letters = cld$Distance_m$Letters,
      Treatment = unique(.$Treatment)
    )
  }) %>%
  ungroup()

letters_within_treatment

letters_within_distance <- quad_change %>%
  group_by(Distance_m) %>%
  do({
    model <- aov(log_urchin_change ~ Treatment, data = .)
    tukey <- TukeyHSD(model)
    cld <- multcompLetters4(model, tukey)
    data.frame(
      Treatment = names(cld$Treatment$Letters),
      Letters = cld$Treatment$Letters,
      Distance_m = unique(.$Distance_m)
    )
  }) %>%
  ungroup()

letters_within_distance
```

```{r}
# Ensure consistent type across data frames
distance_urchin <- distance_urchin %>%
  mutate(Distance_m = as.character(Distance_m))

letters_within_treatment <- letters_within_treatment %>%
  mutate(Distance_m = as.character(Distance_m))

letters_within_distance <- letters_within_distance %>%
  mutate(Distance_m = as.character(Distance_m))

distance_urchin_cld <- distance_urchin %>%
  left_join(letters_within_treatment, by = c("Treatment", "Distance_m")) %>%
  left_join(letters_within_distance, by = c("Treatment", "Distance_m"), suffix = c("_within_treat", "_within_dist"))
```

```{r}
ggplot(distance_urchin_cld, aes(x = Distance_m, y = mean, color = Treatment)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.3) +
  geom_line(aes(group = Treatment)) +
  geom_text(aes(label = Letters_within_treat), 
            vjust = -1.2,
            hjust = 1,
            size = 5, 
            fontface = "bold") +
  labs(
    title = "Change in Total Sea Urchin Density (# per m²)",
    x = "Distance from GA Bar (m)",
    y = "Log10 Fold Change (24 hr / 0 hr)",
    color = "Treatment"
  )+
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 15, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
  
```



#urchin densities by location
```{r}
library(sf)
library(dplyr)

# Purple urchins
purple_sf <- GA %>%
  group_by(Month, Plot, ID, DRIFT.ADDED) %>%
  summarise(
    Density = mean(PURPLE.SEA.URCHIN, na.rm = TRUE),
    LAT = mean(LAT, na.rm = TRUE),
    LONG = mean(LONG, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  st_as_sf(coords = c("LONG", "LAT"), crs = 4326)  # WGS84

red_sf <- GA %>%
  group_by(Month, Plot, ID, DRIFT.ADDED) %>%
  summarise(
    Density = mean(RED.SEA.URCHIN, na.rm = TRUE),
    LAT = mean(LAT, na.rm = TRUE),
    LONG = mean(LONG, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  st_as_sf(coords = c("LONG", "LAT"), crs = 4326)

library(ggplot2)

ggplot(purple_sf) +
  geom_sf(aes(color = DRIFT.ADDED, size = Density), alpha = 0.8) +
  facet_wrap(~Month, nrow = 1) +
  scale_color_manual(values = c("NO" = "gray", "YES" = "forestgreen"), name = "Drift Added") +
  scale_size_continuous(range = c(1, 5), name = "Density (urchins/m²)") +
  labs(
    title = "Purple Urchin Density by ARKEV",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_classic(base_size = 14)
```

```{r}
library(ggplot2)

plot_urchins <- function(sf_data, month_name, species_name) {
  month_data <- sf_data %>% filter(Month == month_name)
  
  ggplot(month_data) +
    geom_sf(aes(color = DRIFT.ADDED, size = Density), alpha = 0.8) +
    scale_color_manual(values = c("NO" = "gray", "YES" = "forestgreen"), name = "Drift Added") +
    scale_size_continuous(range = c(3, 7), name = "Density (urchins/m²)") +
    labs(
      title = paste(species_name, "Density —", month_name),
      x = "Longitude",
      y = "Latitude"
    ) +
    theme_classic(base_size = 14)
}

months <- unique(purple_sf$Month)
purple_plots <- lapply(months, function(m) plot_urchins(purple_sf, m, "Purple Urchin"))
print(purple_plots)
```




#Kelp weight
```{r}
kelp_weight = read.csv("GA_kelp_weights.csv", header=TRUE)

kelp_weight <- kelp_weight %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))%>%
  mutate(Month = month(Date, label = TRUE, abbr = TRUE))%>%
  mutate(Kelp.weight = as.numeric(Kelp.weight))
str(kelp_weight)

KW_noaug <- kelp_weight %>%
  filter(Month != "Aug")

str(KW_noaug)
```
```{r}
unique(kelp_weight$Time)

kelp_weight %>%
  count(Month, ARKEV, Drift, Time) %>%
  pivot_wider(names_from = Time, values_from = n, values_fill = 0)

kelp_weight %>%
  summarise(n = n(), .by = c(Month, ARKEV, Drift, Time)) %>%
  filter(n > 1)
```



```{r}
#calculate percent change to standardize across months

#calc mean and se weight for month
weight_change_month <- kelp_weight %>%
  group_by(Month, ARKEV, Drift, Time) %>%
  summarise(
    mean_weight = mean(Kelp.weight, na.rm = TRUE),
    se_weight = sd(Kelp.weight, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = Time, values_from = c(mean_weight, se_weight)) %>%
  mutate(
    Treatment = paste(Drift, ARKEV, sep = " & "),
    # Percent change: ((Post - Pre) / Pre) * 100
    weight_change = ((mean_weight_Post - mean_weight_Pre) / mean_weight_Pre) * 100,
    # Approximate SE propagation (in % units)
    weight_se = sqrt(se_weight_Post^2 + se_weight_Pre^2) / mean_weight_Pre * 100
  )


#Calculate overall
weight_change <- kelp_weight %>%
  group_by(ARKEV, Drift, Time) %>%
  summarise(
    mean_weight = mean(Kelp.weight, na.rm = TRUE),
    se_weight = sd(Kelp.weight, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = Time, values_from = c(mean_weight, se_weight)) %>%
  mutate(
    Treatment = paste(Drift, ARKEV, sep = " & "),
    weight_change = ((mean_weight_Post - mean_weight_Pre) / mean_weight_Pre) * 100,
    weight_se = sqrt(se_weight_Post^2 + se_weight_Pre^2) / mean_weight_Pre * 100
  )

```

```{r}
#plot
ggplot(weight_change_month, aes(x = Treatment, y = weight_change, fill = Drift)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = weight_change - weight_se, ymax = weight_change + weight_se),
                width = 0.2, position = position_dodge(0.9)) +
  facet_wrap(~Month) +
  labs(title = "24 hr Change in Grazer Assay Kelp Weight (g)",
       x = "Treatment",
       y = "Percent Change in Kelp Weight (g)"
       ) +
  scale_x_discrete(limits = c("NO_DRIFT & Control", "NO_DRIFT & ARKEV", "DRIFT & ARKEV")) + 
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", angle = 60, hjust = 0.5, vjust= 0.5),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.position = "top"
  )


ggplot(weight_change, aes(x = Treatment, y = weight_change)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = weight_change - weight_se, ymax = weight_change + weight_se),
                width = 0.2, position = position_dodge(0.9)) +
  labs(title = "24 hr Change in Grazer Assay Kelp Weight (g)",
       x = "Treatment",
       y = "Percent Change in Kelp Weight (g)"
       ) +
   scale_x_discrete(limits = c("NO_DRIFT & Control", "NO_DRIFT & ARKEV", "DRIFT & ARKEV")) + 
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", hjust = 0.5, vjust= 0.5),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.position = "top"
  )

```

```{r}
library(car)
#run anova
ga_month_lm  <- lm(Kelp.weight ~ Month * Drift, data = kelp_weight)
Anova(ga_month_lm, type = 3)

#run anova
ga_aov <- aov(Kelp.weight ~ ARKEV * Drift, data = kelp_weight)

# Summary of the ANOVA result
summary(ga_aov)
```

```{r}
#run anova without august
ga_noaug_lm <- lm(Kelp.weight ~ ARKEV * Drift, data = KW_noaug)
Anova(ga_noaug_lm)
```

```{r}
#calculate percent change without august

noaug_weight_change <- KW_noaug %>%
  group_by(ARKEV, Drift, Time) %>%
  summarise(
    mean_weight = mean(Kelp.weight, na.rm = TRUE),
    se_weight = sd(Kelp.weight, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = Time, values_from = c(mean_weight, se_weight)) %>%
  mutate(Treatment = paste(Drift, ARKEV, sep = " & "))

# Calculate percent change and propagated SE
noaug_weight_change <- noaug_weight_change %>%
  mutate(
    weight_change = ((mean_weight_Post - mean_weight_Pre) / mean_weight_Pre) * 100,
    weight_se = sqrt(se_weight_Post^2 + se_weight_Pre^2) / mean_weight_Pre * 100
  )

str(noaug_weight_change)
```

```{r}
ggplot(noaug_weight_change, aes(x = Treatment, y = weight_change)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = weight_change - weight_se, ymax = weight_change + weight_se),
                width = 0.2, position = position_dodge(0.9)) +
  labs(
    title = " Change in Grazer Assay Kelp Weight (24 hr)",
    x = "",
    y = "Mean Change in Kelp Weight (%)"
  ) +
  scale_x_discrete(limits = c("NO_DRIFT & Control", "NO_DRIFT & ARKEV", "DRIFT & ARKEV")) + 
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", angle = 15, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.position = "top"
  )
```


```{r}
KW_noaug_summary <- KW_noaug %>%
  group_by(Plot, ID, ARKEV, Drift, Time) %>%
  summarise(mean_weight = mean(Kelp.weight, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Time, values_from = mean_weight) %>%
  mutate(weight_change = ((Post - Pre) / Pre) * 100,
         Treatment = paste(Drift, ARKEV, sep = " & ")) %>%
  drop_na(weight_change)
str(KW_noaug_summary)

anova_drift <- aov(weight_change ~ Drift, data = KW_noaug_summary)
summary(anova_drift)

anova_treatment <- aov(weight_change ~ Treatment, data = KW_noaug_summary)
summary(anova_treatment)
TukeyHSD(anova_treatment)

```

```{r}
library(multcompView)

# Extract Tukey results
tukey_drift <- TukeyHSD(anova_drift)

# Generate compact letter display
cld_drift <- multcompLetters4(anova_drift, tukey_drift)

# Create a clean letters data frame
drift_letters_df <- data.frame(
  Drift = names(cld_drift$Drift$Letters),
  Letters = cld_drift$Drift$Letters
)

# Merge letters into the summary data
KW_noaug_summary_letters <- KW_noaug_summary %>%
  group_by(Drift) %>%
  summarise(
    mean_change = mean(weight_change, na.rm = TRUE),
    se_change = sd(weight_change, na.rm = TRUE) / sqrt(n())
  ) %>%
  left_join(drift_letters_df, by = "Drift")
```

```{r}
ggplot(KW_noaug_summary_letters, aes(x = Drift, y = mean_change)) +
  geom_bar(stat = "identity", fill = "forestgreen", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_change - se_change, ymax = mean_change + se_change),
                width = 0.2) +
  geom_text(aes(label = Letters, y = mean_change + se_change + 5),
            size = 6, fontface = "bold") +
  theme_classic() +
  labs(
    title = "Change in Grazer Assay Kelp Weight by Drift Treatment",
    x = "",
    y = "Mean Change in Kelp Weight (%)"
  ) +
  theme(
    axis.text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 15, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )
```

```{r}
# Plot with jittered points
ggplot() +
  # Individual points (raw data)
  geom_jitter(
    data = KW_noaug_summary,
    aes(x = Drift, y = weight_change),
    width = 0.1, size = 3, alpha = 0.6, color = "grey30"
  ) +
  # Bars for mean ± SE
  geom_bar(
    data = KW_noaug_summary_letters,
    aes(x = Drift, y = mean_change),
    stat = "identity", fill = "steelblue", alpha = 0.7, width = 0.6
  ) +
  geom_errorbar(
    data = KW_noaug_summary_letters,
    aes(x = Drift,
        ymin = mean_change - se_change,
        ymax = mean_change + se_change),
    width = 0.15
  ) +
  # Letters above error bars
  geom_text(
    data = KW_noaug_summary_letters,
    aes(x = Drift, y = mean_change + se_change + 5, label = Letters),
    size = 6, fontface = "bold"
  ) +
  labs(
    title = "Change in Kelp Weight by Drift Treatment",
    x = "",
    y = "Mean Change in Kelp Weight (%)"
  ) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 15, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )
```
