---
title: "kelp_on_arkev_2026"
output: html_document
date: "2026-05-08"
---

```{r}
#installing packages
library(dplyr)
library(tidyr)
library(lubridate)
library(googlesheets4)
library(purrr) 
```

```{r}
#load kelp on arkev data
koa_master = read_sheet("https://docs.google.com/spreadsheets/d/1Xk0a_PX9X-BuZQ9ZkRx0DFtRe8qcjAYHlaMz7DiH1ws/edit?gid=0#gid=0") %>%
 filter(!is.na(kelp_on_ARKEV)) |>
  mutate(stipe_length = as.numeric(stipe_length),
         kelp_on_ARKEV = as.numeric(kelp_on_ARKEV),
         reproductive_kelp_on_ARKEV = as.numeric(reproductive_kelp_on_ARKEV),
         blade_number = as.numeric(blade_number),
         reproductive_blade_number = as.numeric(reproductive_blade_number),
         days_since_outplanting = as.numeric(days_since_outplanting))
str(koa_master)
```

#noyo kelp on arkev
```{r}
library(stringr)
library(forcats)
koa_noyo <- koa_master |>
  mutate(
    date = as.Date(paste(year, month, day, sep = "-"), format = "%Y-%m-%d")  # fix here
  ) |>
  distinct(date, ARKEV_unit, .keep_all = TRUE) |>
  mutate(plot = str_extract(ARKEV_unit, "\\d+")) |>
  mutate(plot = fct_relevel(ARKEV_type,
                            "PVC",
                            "WOOD"
                            )) |>
    group_by(date) |>
  mutate(total = sum(kelp_on_ARKEV)) |>
  group_by(date, ARKEV_type) |>
  mutate(total_type = sum(kelp_on_ARKEV)) |>
  group_by(date, ARKEV_type) |>
  mutate(mean = mean(kelp_on_ARKEV),
         se = sd(kelp_on_ARKEV)/sqrt(length(kelp_on_ARKEV))) |>
  filter(site == "Noyo")


```


```{r}
library(ggplot2)
ggplot(koa_noyo, aes(x= date, y= kelp_on_ARKEV, color = ARKEV_type, group = ARKEV_unit)) +
  geom_point(size=2.5) +
  geom_line(size=1) +
  labs(
    title = "Number of Spliced Sporophytes per ARKEV",
    x = "Month",
    y = "Total # Spliced Individuals"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 15, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.title = element_text(size = 13, face = "bold"),
    legend.text = element_text(size = 13),
    legend.position = "right"
  )

```


```{r}
ggplot(koa_noyo, aes(x= date, y= total_type, color = ARKEV_type)) +
  geom_point(size=2.5) +
  geom_line(size=1) +
  labs(
    title = "Number of Spliced Sporophytes by ARKEV",
    x = "Month",
    y = "Total # Spliced Individuals"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 15, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.title = element_text(size = 13, face = "bold"),
    legend.text = element_text(size = 13),
    legend.position = "right"
  )
```


```{r}
ggplot(koa_noyo, aes(x= date, y= mean, color = ARKEV_type)) +
  geom_point(size=2.5) +
  geom_line(size=1) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, size=0.5) +
  labs(
    title = "Number of Spliced Sporophytes by ARKEV",
    x = "Month",
    y = "Total # Spliced Individuals"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 15, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    title = element_text(size = 15, face = "bold"),
    legend.title = element_text(size = 13, face = "bold"),
    legend.text = element_text(size = 13),
    legend.position = "right"
  )

```

```{r}
stipe_noyo = koa_master %>%
    mutate(
    date = as.Date(paste(year, month, day, sep = "-"), format = "%Y-%m-%d")  # fix here
  ) |>
  filter(!is.na(stipe_length)) |>
  group_by(date) |>
  mutate(avg_length = mean(stipe_length),
            se = sd(stipe_length)/sqrt(length(stipe_length)),
         max = max(stipe_length)) |>
  group_by(date) %>%
  mutate(avg_length = mean(stipe_length),
            se_length = sd(stipe_length)/sqrt(length(stipe_length)),
         max_length = max(stipe_length)) %>%
group_by(date, ARKEV_type) |>
  mutate(avg_length_type = mean(stipe_length),
            se_type = sd(stipe_length)/sqrt(length(stipe_length)),
         max_type = max(stipe_length)) 
  
```

```{r}



```{r}
koa_1 <- koa_noyo %>%
  filter(date == "2026-05-25") %>%
  dplyr::select(date,site,ARKEV_type, ARKEV_unit,kelp_on_ARKEV)

koa_1 <- data.frame(
  ARKEV_type = c(rep("PVC", 40), rep("WOOD", 10)),
  kelp_on_ARKEV = c(
    9, 10, 9, 8, 4, 8, 7, 9, 6, 6, 6, 8, 9, 9, 8, 9, 9, 10, 8, 9,
    7, 4, 9, 8, 10, 7, 5, 7, 8, 8, 6, 7, 6, 6, 3, 7, 8, 5, 8, 7,  # PVC
    1, 4, 1, 4, 0, 6, 0, 0, 0, 1))                                    # WOOD

all_x        <- 0:10
pvc_heights  <- sapply(all_x, function(k) sum(koa_1$kelp_on_ARKEV == k & koa_1$ARKEV_type == "PVC"))
wood_heights <- sapply(all_x, function(k) sum(koa_1$kelp_on_ARKEV == k & koa_1$ARKEV_type == "WOOD"))

stacked <- rbind(PVC = pvc_heights, WOOD = wood_heights)
totals  <- colSums(stacked)

max_y <- max(colSums(stacked))
y_top <- ceiling(max_y / 2) * 2 + 2

par(mar = c(4.5, 4.5, 3, 1), bg = "white")

bp <- barplot(stacked,
  names.arg  = all_x,
  col        = c("#2E7D6E", "#C4622D"),
  border     = "black",
  space      = 0.15,
  ylim       = c(0, 11),
  xlab       = "Kelp count per ARKEV unit",
  ylab       = "Number of ARKEV units",
  main       = "Kelp on ARKEV — 5/25/26 Noyo",
  cex.axis   = 0.9,
  cex.names  = 0.9,
  cex.lab    = 1,
  las        = 1,
  legend.text = rownames(stacked),
  args.legend = list(x = "topleft", bty = "n", cex = 0.9,
                     fill = c("#C4622D","#2E7D6E"), border = NA))
text(x      = bp,
     y      = totals + 0.3,
     labels = ifelse(totals > 0, totals, ""),
     cex    = 0.85,
     font   = 2,
     adj    = c(0.5, 0))
 
```

